# Preliminaries

```{r setup}
library(magrittr)
library(lubridate)
library(profvis)
```

# Hello World

It's time to learn about coding in R!

## 18 Pipes and 19 Functions

```{r}
rescale01 <- function(x) {
    rng <- range(x, na.rm = TRUE)
    (x - rng[1]) / (rng[2] - rng[1])
}

rescale_negative <- function(x) {
    rng <- range(x, na.rm = TRUE)
    2 * (x - rng[1]) / (rng[2] - rng[1]) - 1
}

square <- function(x) {
    x * x
}

test <- c(0, 4, 5, 2, Inf)

# rescale_negative(test) %>%
#     square %>%
#     rescale01
rescale01(test)
```

## 19 Functions Challenge

Write your own functions to compute the variance and skewness of a numeric 
vector. Variance is defined as 

$Var(x) = \frac{1}{n-1} \Sigma_{i=1}^n (x_i - m)^2$

where $m$ is the sample mean (I'm lazy and don't know how to do the $x$ with a 
bar). Skewness is defined as 

$Skew(x) = \frac{\frac{1}{n-2} \left(\Sigma_{i=1}{n}(x_i - m)^3\right)}{Var(x)^{3/2}}$

```{r}
total_mean_diff <- function(x, power) {
    avg <- mean(x)
    sum((x - avg) ^ power)
}

variance <- function(x) {
    total_mean_diff(x, 2) / (length(x) - 1)
}

skew <- function(x) {
    total_mean_diff(x, 3) / ((length(x) - 2) * variance(x) ^ (3 / 2))
}

test <- c(0, 2, 4, 5, 10)

variance(test)

skew(test)
```

## 19.3.1 Exercises

```{r}
starts_with <- function(string, prefix) {
    substr(string, 1, nchar(prefix)) == prefix
}

starts_with("asdf", "as")
starts_with("asdf", "ad")

remove_last <- function(x) {
    if (length(x) <= 1) return(NULL)
    x[-length(x)]
}

remove_last(c(1, 2, 3))

rep_like <- function(x, y) {
    rep(y, length.out = length(x))
}

rep_like(c(1, 2, 3), 5)
```

## 19.4.4 Time exercise

Write a greeting function that says "good morning", "good afternoon", or 
"good evening" depending on the time of day. You'll want `lubradiate::now()`.

```{r}
greeting <- function(time = NULL) {
    if (is.null(time)) time <- lubridate::now()
    lubridate::hour(time) %>%
    cut(breaks = c(0, 4, 12, 20, 24),
        labels = c("evening", "morning", "afternoon", "evening")) %>%
    stringr::str_c("good ", .)
}

greeting(now())
greeting(hm("08:12"))
greeting(hm("13:12"))
greeting(hm("23:12"))
```

## 19.4.4 Exercises

```{r}
fizzbuzz <- function(x) {
    stopifnot(length(x) == 1)
    if (x %% 3 == 0) {
        if (x %% 5 == 0) return("fizzbuzz")
        return("fizz")
    } else if (x %% 5 == 0) {
        return("buzz")
    }
    x
}

fizzbuzz(9)
fizzbuzz(c(1, 2))
fizzbuzz(20)
fizzbuzz(30)
fizzbuzz(8)

switch(4,
    a = 1,
    b = 2,
    "s",
    "t"
)

temp_feel <- function(temp) {
    cut(temp, breaks = c(-Inf, 0, 10, 20, 30, Inf),
        labels = c("freezing", "cold", "cool", "warm", "hot"))
}
temp_feel(c(24, 18, -2))
```

```{r}
last_value <- function(x) {
    x[length(x)]
}

last_value(c(1, 5, 1, 34, 2))
```

```{r}
even_pos <- function(x) {
    if (length(x) < 2) return(NULL)
    x[seq(2, length(x), by = 2)]
}

even_pos(c(1, 5, 1, 34, 2))
```

```{r}
evens_only <- function(x) {
    x[x %% 2 == 0] %>%
        na.omit()
}

evens_only(c(1, 2, 3, 4, NA))
```

## 21.2.1 For loop exercises

```{r}
output <- vector("double", ncol(mtcars))
for (i in seq_along(mtcars)) {
    output[[i]] <- mean(mtcars[[i]])
}
output
```

```{r}
for (i in seq_along(mtcars)) {
    print(typeof(mtcars[[i]]))
}
```

```{r}
for (i in seq_along(mtcars)) {
    print(length(mtcars[[i]][!duplicated(mtcars[[i]])]))
}
```

```{r}
for (i in c(-10, 0, 10, 100)) {
    print(rnorm(1, i))
}
```

```{r}
out <- ""
for (x in letters) {
    out <- stringr::str_c(out, x)
}
out

paste(letters, collapse = "")
```

```{r}
x <- sample(100)

sd <- 0
for (i in seq_along(x)) {
    sd <- sd + (x[i] - mean(x)) ^ 2
}
sd <- sqrt(sd / (length(x) - 1))
sd

sqrt(sum((x - mean(x)) ^ 2) / (length(x) - 1))

sd(x)
```

```{r}
x <- runif(100)

out <- vector("numeric", length(x))
out[1] <- x[1]
for (i in 2:length(x)) {
    out[i] <- out[i - 1] + x[i]
}
out

cumsum(x)
```

### Evaluating performance

```{r}
x <- sample(1000, size = 1, replace = TRUE)
```

```{r}
f <- function(x) {
    output <- vector("numeric", length(x))
    for (i in seq_along(x)) {
        output[i] <- x[[i]]
    }
}
```

```{r}
f <- function(x) {
    output <- vector("numeric", 0)
    for (i in seq_along(x)) {
        output <- c(output, x[[i]])
    }
}
```

```{r}
profvis(f(x))
```

## 21.3.5 Exercises

```{r}
out <- vector("list", 0)
x <- sample(100, 3)
for (i in seq_along(x)) {
    out[[i]] <- names(x)[[i]]
}
names(x)
out
```

```{r}
files <- dir("data/", pattern = "\\.csv$", full.names = TRUE)
df <- data.frame()
for (file in files) {
    df <- rbind(df, read.csv(file))
}
df
```

```{r}
library(dplyr)
files <- dir("data/", pattern = "\\.csv$", full.names = TRUE)
l <- vector("list", 0)
for (i in seq_along(files)) {
    l[[i]] <- read.csv(files[[i]])
}
df <- bind_rows(l)
df
```